A Motion Capture and Imitation Learning-based Approach to Robot Control

### 2022-06-26 tex-arxiv/article.tex
conclusions first draft

When it comes to the use of motion capture data, previous work in the robotics and imitation learning corner is quite sparse, possibly due to the costly and specialized nature of the equipment involved. One previously considered direction has been the use of consumer-grade motion tracking equipment for collecting demonstrations \citep{jha2017imitation}. Unlike our work, they employ a single, relatively inexpensive sensor unit to generate the motion tracking data, and the main focus of the work is on accurate extraction of the demonstrations rather than modelling and extrapolation -- which is left as something of an afterthought, with cluster and k-nearest interpolation methods used for inference. Outside the imitation learning space, related work has been done in human motion modelling utilizing RNNs trained on motion capture data \citep{Fragkiadaki_2015_ICCV}. The main difference between research in this direction and our work lies in the fact that we are looking for ways to translate human motion into paths taken by a robot, and do not consider the human kinematic model. Additionally, we place an emphasis on extracting particular implicit signals such as release timing and desired target coordinates.

@@

>>>>

When it comes to the use of motion capture data, previous work in the robotics and imitation learning corner is quite sparse, possibly due to the costly and specialized nature of the equipment involved. One previously considered direction has been the use of consumer-grade motion tracking equipment for collecting demonstrations \citep{jha2017imitation}. Unlike our work, they employ a single, relatively inexpensive sensor unit to generate the motion tracking data, and the main focus of the work is on accurate extraction of the demonstrations rather than modelling and extrapolation -- which is left as something of an afterthought, with cluster and k-nearest interpolation methods used for inference. Others have applied the previously described dynamic motion primitives to this transfer of human motion to robot configuration space \citep{vuga2013mocap}. Outside the imitation learning space, related work has been done in human motion modelling utilizing RNNs trained on motion capture data \citep{Fragkiadaki_2015_ICCV}. The main difference between research in this direction and our work lies in the fact that we are mapping directly to robot kinematics, don't consider the human kinematic model and place an emphasis on extracting additional implicit signals from the observations.

@@

The original goal we set out to accomplish was to devise a framework for recording demonstrations by human actors and using these to execute an object throwing task as a stand-in for a variety of similar future applications. In this, we were mostly successful -- using motion capture as the data collection mechanism, it proved possible to employ a series of analytic pre-processing steps to turn raw recording data into demonstration data sets suitable for training two varieties of artificial neural networks -- feedforward and recurrent. The models were structured to be able to operate either in-the-loop or autoregressively as forward planners, and it was this latter approach that was further explored -- we developed tools for their visualization (useful in hyperparameter discovery) and prototype motion planning software that enable execution on real robots. Model outputs were used to successfully perform throws in the physical world.

The data collection step with motion capture equipment proved to be straightforward and easy to customize for our motivating application. However, it did require some attention to how the physical demonstrations were performed -- with allowances for automatic demarcation of distinct demonstrations. A pre-processing pipeline had to be developed to extract these separate trajectories and only their relevant segments, as well as infer hidden state variables through indirect observation. While the latter are likely going to be specific to each application, the methods herein can be trivially adapted so long as these can be expressed as n-th order derivatives of absolute or relative pose variables of tracked objects in the scene. Admittedly, some work, such as target coordinate extrapolation, was entirely specific to this task.

In contrast to some previous work that has explored motion capture in imitation learning, we place a greater emphasis on using general-purpose machine learning methods -- artificial neural networks -- to generate the trajectories at inference, as opposed to simple neighbor methods \citep{jha2017imitation} or highly domain-specific approaches grounded in the use of dynamic motion primitives \citep{vuga2013mocap}. To more objectively evaluate the performance of the policies obtained with various hyperparameters and on different data sets, a series of quantitative metrics are proposed in this paper. In general, it can be said that, according to these, recurrent networks operating on complete state histories outperform simple deep neural networks operating in a Markovian mode -- which is perhaps unsurprising, given the recent successes achieved in applying general-purpose sequence-to-sequence learning methods to the imitation learning domain \citep{reed2022generalist}. As our model precision comparisons rely on indirect estimates rather than empirical data, it is hard to make a direct comparison with state-of-the-art performers in throwing tasks such as \citep{zeng2020tossingbot} -- but it should be noted that they use learning at a higher level -- outputting parameters that describe each throw -- and do not use model outputs to generate Cartesian motion plans directly like we do. This makes their entire approach inherently more coupled to a single type of task than ours.

Potential directions for further research and engineering work based what we have achieved thus far would be:
\begin{itemize}
	\item replacement of simple feedforward neural networks with generative adversarial models, which some of the best performance for pure Markovian policies \citep{ho2016generative}. This way the need for more compute-intensive sequence-to-sequence models that necessitate keeping track of state histories may be obviated;
	\item adapting the pipeline for other tasks of interest, mainly to explore how easily the data pre-processing pipeline generalizes in practice;
	\item train models to predict velocities as the actions, rather than subsequent states, for integration with servocontrollers;
	\item development of a Cartesian arrival time-based time parametrization tool in ROS, something which is currently lacking in the ecosystem.
\end{itemize}
